Reintroduction of resistant frogs facilitates landscape-scale recovery in the presence of a lethal fungal disease

### 2024-04-18 doc/manuscript/translocation_bioarxiv.tex
Update PNAS template
Add new figure, text edits, and latex files

% Please add a significance statement to explain the relevance of your work
\significancestatement{Understanding how species persist despite accelerating global change is
critical for the conservation of biodiversity. Emerging infectious
diseases can have particularly devastating impacts, and few options
exist to reverse these effects. We used large-scale reintroductions of
disease-resistant individuals in an effort to recover a once-common frog
species driven to near-extinction by a disease that has decimated
amphibian biodiversity. Introduction of resistant frogs allowed
reestablishment of viable populations in the presence of disease. In
addition, resistance may be at least partially the result of natural
selection at specific immune function genes, which show evidence for
selection in recovering populations. The evolution of resistance and
reintroduction of resistant individuals could play an important role in
biodiversity conservation in our rapidly changing world.}

@@

% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
\keywords{}

\begin{abstract}
Vast alteration of the biosphere by humans is causing a sixth mass
extinction, driven in part by an increase in emerging infectious
diseases. The emergence of the lethal fungal pathogen
(\emph{Batrachochytrium dendrobatidis}; ``Bd'') has devastated global
amphibian biodiversity, with hundreds of species experiencing declines
or extinctions. With no broadly applicable methods available to reverse
these impacts in the wild, the future of many amphibians appears grim.
The once-common mountain yellow-legged (MYL) frog is emblematic of
amphibians threatened by Bd. Although most MYL frog populations are
extirpated following disease outbreaks, some persist and eventually
recover. Frogs in these recovering populations have increased resistance
against Bd infection, consistent with evolution of resistant genotypes
and/or acquired immunity. We conducted a 15-year landscape-scale
reintroduction study and show that frogs collected from recovering
populations and reintroduced to vacant habitats can reestablish
populations despite the presence of Bd. In addition, results from
viability modeling suggest that many reintroduced populations have a low
probability of extinction over 50 years. To better understand the role
of evolution in frog resistance, we compared the genomes of MYL frogs
from Bd-naive and recovering populations. We found substantial
differences between these categories, including changes in immune
function loci that may confer increased resistance, consistent with
evolutionary changes in response to Bd exposure. These results provide a
rare example of how reintroduction of resistant individuals can allow
the landscape-scale recovery of disease-impacted species. This example
has broad implications for the many taxa worldwide that are threatened
with extinction by novel pathogens.
\end{abstract}

@@

Evidence of natural recovery in the many Bd-impacted amphibian
populations is surprisingly rare \citep[for notable exceptions,
see][]{scheele2017, voyles2018, knapp2016}, suggesting that
disease-caused declines will be difficult to reverse. This apparent low
resilience to disease effects may be due to the limited ability of many
amphibians to develop Bd resistance and/or tolerance, which in turn,
could also lessen the effectiveness of potential Bd mitigation
strategies. Following pathogen arrival in a host population, resistance
(ability to limit pathogen burden) and tolerance (ability to limit the
harm caused by a particular burden) are key mechanisms to reduce disease
impacts \citep{raberg2009} and facilitate population persistence and
recovery \citep{brannelly2021}. Host immunity and evolution both play
important roles in the development of resistance and tolerance, and
utilizing these factors would seem a promising approach to developing
effective strategies to mitigate disease impacts in the wild
\citep{garner2016, woodhams2011}. However, several aspects of the
amphibian-Bd system present difficult obstacles, including (i) the
general inability of amphibians to mount an effective immune response
against Bd infection \citep{rosenblum2012, fites2013, grogan2018a}, and
(ii) the apparent rarity of evolution of more resistant/tolerant
genotypes \citep[but see][]{savage2016, grogan2018b}. These factors
suggest that reintroduction of amphibians into sites to reestablish
populations extirpated by Bd will often result not in population
recovery, but instead in the rapid reinfection and mortality of the
introduced animals and/or their progeny
\citep{hammond2021, knapp2022, stockwell2008, scheele2021}. If true, the
future of many amphibian species threatened by Bd appears bleak.

@@

Despite the catastrophic impact of Bd on MYL frogs, wherein most
Bd-naive populations are extirpated following Bd arrival
\citep{vredenburg2010}, some populations have persisted after epizootics
\citep[during which Bd infection intensity on frogs is very
high,][]{briggs2010} and are now recovering
(Figure~\ref{fig-recovery-model}) \citep{knapp2016}. Frogs in these
recovering populations show reduced susceptibility to Bd infection
\citep{knapp2016}, with infection intensity (``load'') on adults
consistently in the low-to-moderate range
\citep{briggs2010, knapp2011, joseph2018}. This reduced susceptibility
is evident even under controlled laboratory conditions
\citep{knapp2016}, indicative of host resistance against Bd infection
(and not simply an effect of factors external to individual frogs, e.g.,
environmental conditions). In addition to frogs from recovering
populations having higher resistance to Bd infection than those from
naive populations, they could also have higher tolerance, but no data
are currently available to evaluate this possibility. Therefore, we
focus on resistance throughout this paper. The observed resistance of
MYL frogs could be the result of several non-mutually exclusive
mechanisms, including natural selection for more resistant genotypes
\citep{savage2016, grogan2018b}, acquired immunity \citep{grogan2018a},
and/or inherent between-population differences that pre-date Bd
exposure. The possible evolution of MYL frog resistance and subsequent
population recovery is consistent with that expected under
``evolutionary rescue'', whereby rapid evolutionary change increases the
frequency of adaptive alleles and restores positive population growth
\citep{carlson2014, searle2020}. This intriguing possibility also
suggests an opportunity to expand recovery beyond the spatial scale
possible under natural recovery by utilizing resistant frogs from
recovering populations in reintroductions to vacant habitats
(Figure~\ref{fig-recovery-model}) \citep{joseph2018, mendelson2019}.

In the current study, we had three primary objectives. First, to
determine whether the reintroduction of resistant MYL frogs obtained
from populations recovering from Bd-caused declines allows the
successful reestablishment of extirpated populations despite ongoing
disease, we conducted a 15-year landscape-scale frog reintroduction
effort (Figure~\ref{fig-recovery-model}). Second, to extend our
inferences of population recovery well beyond the temporal extent of our
reintroduction study, we developed a model to estimate the probability
of persistence for the reintroduced populations over a multi-decadal
period (Figure~\ref{fig-recovery-model}). Third, given the importance of
resistance for frog survival, population establishment, and long-term
viability (this study), we conducted a genomic study using exome capture
methods to determine whether MYL frogs in recovering populations show
evidence of selection and whether these genomic changes are associated
with resistance (Figure~\ref{fig-recovery-model}).

@@

To determine whether MYL frogs from recovering populations can be used
to reestablish extirpated populations, we conducted 24 reintroductions
in Yosemite National Park (2006--2020). Each of the reintroductions
involved collection of adult frogs from 1 of 3 recovering, Bd-positive
``donor'' populations and translocating them to 1 of 12 nearby recipient
sites (Figure S1). The donor populations included 2 of
the 5 recovering populations used in the frog evolution study referenced
above (Figure~\ref{fig-allelefrequencies}: population 1 and 4), and
these donor populations contributed frogs for 20 of the 24
translocations. Following translocation, we estimated adult survival and
recruitment of new adults from capture-mark-recapture (CMR) surveys and
obtained counts of tadpoles and juveniles from visual encounter surveys
(VES). Across all translocation sites, the duration of survey time
series was 1--16 years (median = 5).

@@

In summary, results from our frog translocation study indicate that
translocations resulted in (i) relatively high 1-year survival of
translocated adults, as well as reproduction and recruitment, at the
majority of recipient sites, (ii) 1-year survival of adults is
influenced by site characteristics, weather conditions, and donor
population (but not Bd load), and (iii) based on the relatively small
changes in Bd load after translocation, loads appear more strongly
influenced by frog characteristics (e.g., resistance) than site
characteristics. Together, these results indicate that frogs
translocated from recovering populations can maintain the benefits of
resistance in non-natal habitats. In addition, in 3 locations where
longer CMR time series allowed us to assess the survival of new adults
recruited to the population, naturally-recruited adults had equivalent
or higher survival probabilities than the originally translocated adults
(Figure S4). This suggests that frog
resistance is maintained across generations. All of the conditions
described above are supportive of population establishment and long-term
population growth.

@@

\subsection*{Frog evolution in response to Bd}

Results from the preceding sections indicate the critical role of frog
resistance in post-translocation frog survival, population growth, and
population viability. As such, identifying the mechanisms underlying
this resistance would fill a key gap in our understanding of the factors
that promote population resilience in the presence of disease. Although
natural selection for more resistant frog genotypes, and evolutionary
rescue, may be foundational to the ability of frogs to recover despite
ongoing Bd infection, for MYL frogs the role of disease-mediated
selection in these processes remains unknown.

To determine whether MYL frog populations show genomic patterns
consistent with an evolutionary response to Bd, we compared frog exomes
(i.e., coding region of a genome) between populations with contrasting
histories of Bd exposure. Specifically, we compared frog genomes sampled
in 4 populations that have not yet experienced a Bd-caused epizootic
(``naive'') \citep{zhou2015} versus in 5 populations that experienced a
Bd epizootic during the past several decades and have since recovered to
varying degrees (``recovering''; Figure~\ref{fig-allelefrequencies})
\citep{knapp2016, vredenburg2010}. Bd-exposure histories of the 9 study
populations are based on 10-20 years of VES and Bd surveillance using
skin swabbing \citep[e.g.,][]{knapp2016, zhou2015, wilber2022}. Naive
populations are characterized by large numbers of adults (i.e, typically
1000s), Bd prevalence that is generally 0\% except during occasional Bd
failed invasions \citep[during which Bd loads remain very
low,][]{wilber2022}, and no history of Bd epizootics since we first
surveyed these populations in the late 1990s and early 2000s
\citep{zhou2015}. In contrast, recovering populations exist in an
enzootic state \citep{briggs2010}, characterized by smaller numbers of
adults (generally \textless{} 500), high Bd prevalence \citep[often
\textgreater{} 80\%,][]{knapp2011}, and, in adults, moderate Bd loads
that are typically well below the level expected to cause mortality
\citep{vredenburg2010}. Naive and recovering populations can be
identified unambiguously using these differences in Bd prevalence and
load. Finally, there is no potential for frog dispersal between the 9
study populations due to intervening distances and topography, as well
as the presence of introduced (predatory) fish and fish-induced habitat
fragmentation. (See \textbf{Supporting Information - Frog evolution in
response to Bd - Study design} for additional details regarding the
study design.)

We conducted a principal component analysis (PCA) of the genomic data to
describe the relationships between sampled populations, and then used
two complementary approaches to identify regions of the genome that
differed between naive and recovering populations (i.e., regions under
selection). First, we used a multivariate linear mixed model to evaluate
associations between population type (i.e., naive versus recovering) and
individual variants, including single nucleotide polymorphisms (SNPs)
and insertions/deletions (INDELS), while accounting for population
structure. Second, we used a splined window analysis to identify larger
genomic regions showing differences between population types in
\emph{F\textsubscript{ST}} and nucleotide diversity (\(\pi_{diff}\) =
\(\pi_{naive}\) - \(\pi_{recovering}\)).

@@

Results from the individual variant and splined window analyses show
that recovering populations have signatures of selection on multiple
regions of the genome. The analysis of individual variants identified 11
``outlier'' SNPs (i.e., showing significantly different allele
frequencies between naive versus recovering populations) from 7 distinct
genes across 4 contigs (Figure S6B, C). One
of the 7 identified genes (LOC108802036) does not have an associated
annotation. For the outlier SNPs, frequency differences between the
naive and recovering populations ranged from 0.41 to 0.86. Most of these
SNPs showed frequency differences in only a subset of the sampled
populations (Figure~\ref{fig-allelefrequencies}A, B), but the SNP in
the RIN3 gene showed consistent differences in frequencies across all
populations (Figure~\ref{fig-allelefrequencies}C). This is suggestive
of parallel selection at this locus across multiple populations. The
other 6 outlier variants showed less consistent frequency differences
across the study populations, but for these we still found a
statistically significant signal of selection in 2 of the 3 genetic
clusters (containing populations 5--9;
Figure~\ref{fig-allelefrequencies}, Figure S6A). Therefore, although some outlier variant associations have a more
limited geographic extent than RIN3, they still describe results that
suggest parallel evolutionary changes following Bd exposure.

@@

In summary, our genomic results indicate that the exomes of frogs from
naive and recovering populations show substantial differences,
consistent with parallel evolutionary changes following Bd exposure. The
regions under selection contain several immunologically-relevant genes
and gene families that are directly linked to disease resistance in
other taxa.

@@

Results from our genomic analyses suggest that natural selection for
adaptive alleles is at least partially responsible for the increased
resistance of frogs in recovering populations. We identified multiple
specific alleles and genomic regions showing signatures of selection
between adjacent naive and recovering MYL frog populations, consistent
with selection following Bd exposure. These analysis are based on
samples collected from virtually all of the MYL frog populations
remaining in a naive state, as well as adjacent recovering populations.
This study design produced genetic clusters that each contained at least
one naive and one recovering population, allowing us to detect selection
without the confounding effects of population structure. In addition, we
did not find a reduction in overall genetic variation in the recovering
populations, suggesting that despite localized selection in the genome,
these populations retain adequate genetic diversity for long-term
persistence.

Importantly, some genomic regions that we identified as under selection
are associated with cellular and immunological mechanisms known to
contribute to disease resistance, including in amphibians
\citep{zamudio2020}. For example, the MHC plays an important role in
immunity. In our study, we identified a region that shows evidence of
selection in recovering populations and contains eight genes associated
with either the MHC Class I or Class III regions. These results
corroborate numerous previous studies linking MHC genes to amphibian
resistance against Bd \citep[e.g.,][]{savage2011, bataille2015}.
Similarly, the region with the strongest indication of directional
selection (as measured by \(\pi_{diff}\)) contains the
interferon-related gene GVINP1. Several previous studies of amphibians
have found this gene to be differentially expressed during Bd infection
\citep[e.g.,][]{grogan2018b, ellison2014} and in populations differing
in Bd susceptibility \citep{grogan2018b}. This gene is also strongly
linked to disease in salmon, explaining a notable 20\% of the resistance
phenotype \citep{robledo2020, robledo2018}. We also identified a region,
characterized by high \emph{F\textsubscript{ST}} and low \(\pi_{diff}\),
that contained the complement genes C6 and C7. The complement system
plays an important role in innate immunity \citep{rieraromo2016}, and
our results could indicate that balancing selection is acting in this
region of the genome to favor a diverse set of alleles, as is known for
C6 in humans \citep{soejima2005}. Based on the analysis of individual
outlier variants, the RIN3 gene showed a consistent pattern of allele
frequency differences across all nine of the frog populations sampled in
this study, indicating consistent selection in populations distributed
across a wide geographic area. This gene is associated with immune
response and in \emph{Xenopus} is expressed during appendage
regeneration \citep{fukazawa2009}. Finally, the outlier variant with the
lowest p-value was the uncharacterized gene LOC108802036. In the genome
of another frog species, this gene is located adjacent to a type I
interferon gene (Np-IFNi2) \citep{gan2018}, and together with GVINP1
further suggests the importance of interferon-related genes in this
system. Collectively, the genes associated with these genomic
differences may confer at least some degree of resistance against Bd
infection, an attribute that may be critically important to population
reestablishment and recovery in the presence of Bd.

@@

@@

After parameterizing our model with CMR-estimated adult frog survival
probabilities and other known vital rates
(Table S1), we performed four analyses. First, we
compute the long-run growth rate \(\lambda\) for each of our 12
translocated populations to determine if the populations were
deterministically predicted to grow or decline in the long-run. Second,
we compute the elasticity of \(\lambda\) to four key model parameters to
quantify how much changes in these parameters affected the long-run
growth rate (Figure S5). This also helped us
determine where in the model environmental variation in juvenile
survival and recruitment would have the largest effects on population
dynamics. Third, we included demographic stochasticity and environmental
stochasticity in \(\sigma_{J_1}\) in our model and simulated the 50-year
viability (i.e., 1 - extinction probability) of populations given an
introduction of 40 adult individuals into an unoccupied habitat.
Finally, we fit our model to our longest translocation trajectory to
confirm that our model could reasonably reproduce the observed recovery
trajectories of MYL frogs following reintroductions. Additional details
are provided in \textbf{Supporting Information - Population viability
modeling - Model analysis and simulation}. Code to replicate the
analyses can be found at \url{https://github.com/SNARL1/translocation}.

\hypertarget{frog-evolution-in-response-to-bd-1}{%
\section*{Frog evolution in response to
Bd}\label{frog-evolution-in-response-to-bd-1}}

@@

\caption{\label{fig-recovery-model}For MYL frogs, a conceptual model
depicting the Bd-caused decline and subsequent natural recovery (black
text), facilitated recovery via reintroductions, and the linkages
between these two pathways. Rectangles and hexagons represent outcomes
and processes, respectively. Blue text indicates components that are
included in the current study. The timeline shows the general sequence
of the components, with the dotted portion indicating a projection into
the future.}

>>>>

% Please add a significance statement to explain the relevance of your work
\significancestatement{Understanding how species persist despite accelerating global change is
critical for the conservation of biodiversity. We used reintroductions
of disease-resistant individuals in an effort to recover a once-common
frog species driven to near-extinction by a novel disease that has
decimated amphibians worldwide. Reintroductions allowed establishment of
viable populations in the presence of disease. In addition, resistance
may be at least partially the result of natural selection at specific
immune function genes. The evolution of resistance and reintroduction of
resistant individuals could play an important role in biodiversity
conservation in our rapidly changing world.}

@@

% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
\keywords{evolution | resistance | amphibians | population recovery | reintroduction | disease | \emph{Batrachochytrium dendrobatidis}}

\begin{abstract}
Vast alteration of the biosphere by humans is causing a sixth mass
extinction, driven in part by an increase in infectious diseases. The
emergence of the lethal fungal pathogen (\emph{Batrachochytrium
dendrobatidis}; ``Bd'') has devastated global amphibian biodiversity,
with hundreds of species experiencing declines or extinctions. Lacking
any broadly applicable methods to reverse these impacts, the future of
many amphibians appears grim. The once-common mountain yellow-legged
(MYL) frog is emblematic of amphibians threatened by Bd. Although most
MYL frog populations are extirpated following disease outbreaks, some
persist and eventually recover. Frogs in these recovering populations
have increased resistance against Bd infection, consistent with
evolution of resistant genotypes. We conducted a 15-year landscape-scale
reintroduction study and show that frogs collected from recovering
populations and reintroduced to vacant habitats can reestablish
populations despite the presence of Bd. In addition, the likelihood of
establishment is influenced by site, population, and frog attributes.
Results from viability modeling suggest that many reintroduced
populations have a low probability of extinction over 50 years. To
better understand the role of evolution in frog resistance, we compared
the genomes of MYL frogs from Bd-naive and recovering populations. We
found numerous differences between these categories, including in immune
function loci associated with disease resistance, consistent with
evolutionary changes in response to Bd exposure. These results provide a
rare example of how reintroduction of resistant individuals can allow
the landscape-scale recovery of disease-impacted species, and have broad
implications for the many taxa worldwide that are threatened with
extinction by novel pathogens.
\end{abstract}

@@

Evidence of natural recovery in the many Bd-impacted amphibian
populations is surprisingly rare \citep[for notable exceptions,
see][]{scheele2017, voyles2018, knapp2016}, suggesting that
disease-caused declines will be difficult to reverse. This apparent low
resilience to disease effects may be due to the limited ability of many
amphibians to develop Bd resistance and/or tolerance, which in turn,
could also lessen the effectiveness of potential Bd mitigation
strategies. Following pathogen arrival in a host population, resistance
(ability to limit pathogen burden) and tolerance (ability to limit the
harm caused by a particular burden) are key mechanisms to reduce disease
impacts \citep{raberg2009} and facilitate population persistence and
recovery \citep{brannelly2021}. Host immunity and evolution both play
important roles in the development of resistance and tolerance, and an
enhanced understanding of these processes is important for developing
effective strategies to mitigate disease impacts in the wild
\citep{garner2016, woodhams2011}. However, several aspects of the
amphibian-Bd system present difficult obstacles for the development of
mitigation efforts, including (i) the general inability of amphibians to
mount an effective immune response against Bd infection
\citep{rosenblum2012, fites2013, grogan2018a}, and (ii) the apparent
rarity of evolution of more resistant/tolerant genotypes \citep[but
see][]{savage2016, grogan2018b}. These factors suggest that
reintroduction of amphibians into sites to reestablish populations
extirpated by Bd will often result not in population recovery, but
instead in the rapid reinfection and mortality of the introduced animals
and/or their progeny
\citep{hammond2021, knapp2022, stockwell2008, scheele2021}. If true, the
future of many amphibian species threatened by Bd appears bleak.

@@

Despite the catastrophic impact of Bd on MYL frogs, wherein most
Bd-naive populations are extirpated following Bd arrival
\citep{vredenburg2010}, some populations have persisted after epizootics
\citep[during which Bd infection intensity on frogs is very
high,][]{briggs2010} and are now recovering
(Figure~\ref{fig-recovery-model}) \citep{knapp2016}. Frogs in these
recovering populations show reduced susceptibility to Bd infection
\citep{knapp2016}, with infection intensity (``load'') on adults
consistently in the low-to-moderate range
\citep{briggs2010, knapp2011, joseph2018}. This reduced susceptibility
is evident even under controlled laboratory conditions
\citep{knapp2016}, indicative of host resistance against Bd infection
and not simply an effect of factors external to individual frogs (e.g.,
environmental conditions). In addition to frogs from recovering
populations having higher resistance to Bd infection than those from
naive populations, they could also have higher tolerance, but no data
are currently available to evaluate this possibility. Therefore, we
focus on resistance throughout this paper. The observed resistance of
MYL frogs could be the result of several mechanisms that are not
mutually exclusive, including natural selection for more resistant
genotypes \citep{savage2016, grogan2018b}, acquired immunity
\citep{grogan2018a}, and/or inherent between-population differences that
pre-date Bd exposure. The possible evolution of MYL frog resistance and
subsequent population recovery is consistent with that expected under
``evolutionary rescue'', whereby rapid evolutionary change increases the
frequency of adaptive alleles and restores positive population growth
\citep{carlson2014, searle2020}. This intriguing possibility also
suggests an opportunity to expand recovery beyond the spatial scale
possible under natural recovery by using resistant frogs from recovering
populations in reintroductions to vacant habitats
(Figure~\ref{fig-recovery-model}) \citep{joseph2018, mendelson2019}.

In the current study, we had three primary objectives. First, determine
whether the reintroduction of resistant MYL frogs, obtained from
populations recovering from Bd-caused declines, allows the
reestablishment of populations despite ongoing disease
(Figure~\ref{fig-recovery-model}). We accomplished this using a 15-year
large-scale reintroduction study. Second, extend our inferences of
population recovery beyond the temporal extent of our reintroduction
study to estimate the probability of persistence for the reintroduced
populations over a multi-decadal period
(Figure~\ref{fig-recovery-model}). We did this using a stage-structured
matrix model. Third, given the importance of resistance for frog
survival, population establishment, and long-term viability (this
study), determine whether MYL frogs in recovering populations show
genomic patterns consistent with selection for resistance
(Figure~\ref{fig-recovery-model}). We achieved this with a comparative
study using exome capture methods applied to naive and recovering
populations.

@@

To determine whether MYL frogs from recovering populations can be used
to reestablish extirpated populations, we conducted 24 reintroductions
in Yosemite National Park (2006--2020). Each of the reintroductions
involved collection of adult frogs from 1 of 3 recovering, Bd-positive
``donor'' populations and translocating them to 1 of 12 nearby recipient
sites (Figure S1). The 3 donor populations are among
the largest MYL frog populations in Yosemite, and included 2 of the 5
recovering populations used in the frog evolution study referenced above
(see \textbf{Results - Frog evolution in response to Bd:} population 1
and 4). In addition, this subset of 2 donor populations provided frogs
for 20 of the 24 translocations. Following translocation, we estimated
adult survival and recruitment of new adults from capture-mark-recapture
(CMR) surveys and obtained counts of tadpoles and juveniles from visual
encounter surveys (VES). Across all translocation sites, the duration of
survey time series was 1--16 years (median = 5).

@@

In summary, results from our frog translocation study indicate that (i)
translocations produced relatively high 1-year survival of translocated
adults, as well as reproduction and recruitment, at the majority of
recipient sites, (ii) 1-year survival of adults is influenced by site
characteristics, weather conditions, and donor population (but not Bd
load), and (iii) based on the relatively small changes in Bd load after
translocation, loads appear more strongly influenced by frog
characteristics (e.g., resistance) than site characteristics. Together,
these results indicate that frogs translocated from recovering
populations can maintain the benefits of resistance in non-natal
habitats. In addition, in 3 locations where longer CMR time series
allowed us to assess the survival of new adults recruited to the
population, naturally-recruited adults had equivalent or higher survival
probabilities than the originally translocated adults
(Figure S4). This suggests that frog
resistance is maintained across generations. All of the conditions
described above are supportive of population establishment and long-term
population growth.

@@

\hypertarget{frog-genome-evolution}{%
\paragraph{Frog genome evolution}\label{frog-genome-evolution}}

Results from the preceding sections indicate the critical role of
intrinsic factors (such as frog resistance) in post-translocation frog
survival, population growth, and population viability. As such,
identifying the mechanisms underlying resistance would fill a key gap in
our understanding of the factors that promote population resilience in
the presence of disease. Although natural selection for more resistant
frog genotypes, and evolutionary rescue, may be foundational to the
ability of frogs to recover despite ongoing Bd infection, for MYL frogs
the role of disease-mediated selection in these processes remains
unknown.

To determine whether MYL frog populations show genomic patterns
consistent with an evolutionary response to Bd, we compared frog exomes
(i.e., coding region of a genome) between populations with contrasting
histories of Bd exposure. Specifically, we compared frog genomes sampled
in 4 populations that have not yet experienced a Bd-caused epizootic
(``naive'') \citep{zhou2015} versus in 5 populations that experienced a
Bd epizootic during the past several decades and have since recovered to
varying degrees (``recovering''; Figure~\ref{fig-allelefrequencies})
\citep{knapp2016, vredenburg2010}. Bd-exposure histories of the 9 study
populations are based on 10-20 years of VES and Bd surveillance using
skin swabbing \citep[e.g.,][]{knapp2016, zhou2015, wilber2022}. Naive
populations are characterized by large numbers of adults (i.e, typically
1000s), Bd prevalence that is generally 0\% except during occasional Bd
failed invasions \citep[during which Bd loads remain very
low,][]{wilber2022}, and no history of Bd epizootics since we first
surveyed these populations in the late 1990s and early 2000s
\citep{zhou2015}. In contrast, recovering populations exist in an
enzootic state \citep{briggs2010}, characterized by smaller numbers of
adults (generally \textless{} 500), high Bd prevalence \citep[often
\textgreater{} 80\%,][]{knapp2011}, and, in adults, moderate Bd loads
that are typically well below the level expected to cause mortality
\citep{vredenburg2010}. Naive and recovering populations can be
identified unambiguously using these differences in Bd prevalence and
load. Finally, there is no potential for frog dispersal between the 9
study populations due to intervening distances and topography, as well
as the presence of introduced (predatory) fish and fish-induced habitat
fragmentation.

It is important to note that our genomic analysis is necessarily limited
by the small number of naive populations remaining on the landscape
(nearly all of which we sampled as part of the current study) and their
clumped distribution Figure~\ref{fig-allelefrequencies}. These
characteristics of naive populations precluded a replicated analyses
among pairs of geographically proximate naive and recovering
populations. Instead, to obtain adequate power, we conducted a pooled
analysis across population types to identify genomic regions showing
patterns consistent with selection. (See \textbf{Supporting Information
- Frog evolution in response to Bd - Study design} for additional
details regarding the study design.)

As a first step in our pooled analysis, we conducted a principal
component analysis (PCA) of the genomic data to describe the
relationships between sampled populations. We then used two
complementary approaches to identify regions of the genome that differed
between naive and recovering populations (i.e., regions under
selection). First, we used a multivariate linear mixed model to evaluate
associations between population type (i.e., naive versus recovering) and
individual variants, including single nucleotide polymorphisms (SNPs)
and insertions/deletions (INDELS), while accounting for population
structure. Second, we used a splined window analysis to identify larger
genomic regions showing differences between population types in
\emph{F\textsubscript{ST}} and nucleotide diversity (\(\pi_{diff}\) =
\(\pi_{naive}\) - \(\pi_{recovering}\)).

@@

Results from the individual variant and splined window analyses show
that recovering populations show patterns consistent with selection on
multiple regions of the genome. The analysis of individual variants
identified 11 ``outlier'' SNPs (i.e., showing significantly different
allele frequencies between naive versus recovering populations) from 7
distinct genes across 4 contigs (Figure S6B, C
SI). One of the 7 identified genes (LOC108802036) does not have an
associated annotation. For the outlier SNPs, frequency differences
between the naive and recovering populations ranged from 0.41 to 0.86.
Most of these SNPs showed frequency differences in only a subset of the
sampled populations (Figure~\ref{fig-allelefrequencies}A, B), but the
SNP in the RIN3 gene showed consistent differences in frequencies across
all populations (Figure~\ref{fig-allelefrequencies}C). This is
suggestive of parallel selection at this locus across multiple
populations. The other 6 outlier variants showed less consistent
frequency differences across the study populations, but for these we
still found a statistically significant signal of selection in 2 of the
3 genetic clusters (containing populations 5--9;
Figure~\ref{fig-allelefrequencies}, Figure S6A). Therefore, although some outlier variant associations have a more
limited geographic extent than RIN3, they still describe results that
suggest parallel evolutionary changes following Bd exposure.

@@

In summary, our genomic results indicate that frogs from naive and
recovering populations show differences at several immune-function loci,
consistent with selection following Bd exposure.

@@

Our genomics study used a pooled design that compared MYL frogs from
naive and recovering populations distributed across a similar
latitudinal range. The study design produced frog genetic clusters that
each contained at least one naive and one recovering population,
allowing us to infer selection while attempting to minimize the
confounding effects of population structure. We identified multiple
specific alleles and genomic regions showing differences between naive
and recovering MYL frog populations, consistent with selection following
Bd exposure. In addition, we did not find a reduction in overall genetic
variation in recovering populations, suggesting that despite localized
selection in the genome, some of these populations likely retain
adequate genetic diversity for long-term persistence.

Importantly, some genomic regions that we identified as possibly under
selection are associated with cellular and immunological mechanisms
known to contribute to disease resistance, including in amphibians
\citep{zamudio2020}. For example, the MHC plays an important role in
immunity. In our study, we identified a region that shows evidence of
selection in recovering populations and contains eight genes associated
with either the MHC Class I or Class III regions. These results
corroborate numerous previous studies linking MHC genes to amphibian
resistance against Bd \citep[e.g.,][]{savage2011, bataille2015}.
Similarly, the region with the strongest indication of directional
selection (as measured by \(\pi_{diff}\)) contains the
interferon-related gene GVINP1. Several previous studies of amphibians
have found this gene to be differentially expressed during Bd infection
\citep[e.g.,][]{grogan2018b, ellison2014} and in populations differing
in Bd susceptibility \citep{grogan2018b}. This gene is also strongly
linked to disease in salmon, explaining a notable 20\% of the resistance
phenotype \citep[i.e., gill damage and amoebic load as heritable host
resistance traits,][]{robledo2020, robledo2018}. We also identified a
region, characterized by high \emph{F\textsubscript{ST}} and low
\(\pi_{diff}\), that contained the complement genes C6 and C7. The
complement system plays an important role in innate immunity
\citep{rieraromo2016}, and our results could indicate that balancing
selection is acting in this region of the genome to favor a diverse set
of alleles, as is known for C6 in humans \citep{soejima2005}. Based on
the analysis of individual outlier variants, the RIN3 gene showed a
consistent pattern of allele frequency differences across all nine of
the frog populations sampled in this study, indicating consistent
selection in populations distributed across a wide geographic area. This
gene is associated with immune response and in \emph{Xenopus} is
expressed during appendage regeneration \citep{fukazawa2009}. Finally,
the outlier variant with the lowest p-value was the uncharacterized gene
LOC108802036. In the genome of another frog species, this gene is
located adjacent to a type I interferon gene (Np-IFNi2) \citep{gan2018},
and together with GVINP1 further suggests the importance of
interferon-related genes in this system. Collectively, the genes
associated with these genomic differences may confer at least some
degree of resistance against Bd infection, an attribute that may be
critically important to population reestablishment and recovery in the
presence of Bd.

@@

The results of the meta-analysis indicate the important influence of
several variables on adult survival, and by extension on population
viability. Although we assume that the effects of winter severity and
frog sex are direct effects of the variables themselves, the elevation
and donor population variables are likely serving as proxies for a range
of factors. For example, elevation is positively correlated with several
attributes that may influence frog survival, including the presence of
boulder habitat that is important for frog overwintering, abundance of
frog predators including garter snakes, and frog developmental rates.
The latter can influence the duration of frog life stages that have
differential susceptibility to chytridiomycosis (see below). The
duration of the tadpole and juvenile stages increases with elevation,
but the fact that long-run population growth rate is relatively
insensitive to variation in these durations suggests that developmental
rates of early frog life stages have little effect on the predicted
long-term viability of reintroduced populations. Nonetheless, a clearer
understanding of the mechanistic components of the elevation and donor
population variables will require additional research.

@@

After parameterizing our model with CMR-estimated adult frog survival
probabilities and other known vital rates
(Table S1), we performed four analyses. First, we
compute the long-run growth rate \(\lambda\) for each of our 12
translocated populations to determine if the populations were
deterministically predicted to grow or decline in the long-run. Second,
we compute the elasticity and sensitivity of \(\lambda\) to nine model
parameters to quantify how much changes in these parameters affected the
long-run growth rate (Figure S5). This also
helped us determine where in the model environmental variation in
juvenile survival and recruitment would have the largest effects on
population dynamics. Third, we included demographic stochasticity and
environmental stochasticity in \(\sigma_{J_1}\) in our model and
simulated the 50-year viability (i.e., 1 - extinction probability) of
populations given an introduction of 40 adult individuals into an
unoccupied habitat. Finally, we fit our model to our longest
translocation trajectory to confirm that our model could reasonably
reproduce the observed recovery trajectories of MYL frogs following
reintroductions. Additional details are provided in \textbf{Supporting
Information - Population viability modeling - Model analysis and
simulation}. Code to replicate the analyses can be found at
\url{https://github.com/SNARL1/translocation}.

\hypertarget{frog-genome-evolution-1}{%
\section*{Frog genome evolution}\label{frog-genome-evolution-1}}

@@

\caption{\label{fig-recovery-model}For MYL frogs, a conceptual model
depicting the Bd-caused decline and subsequent natural recovery (black
text), facilitated recovery via reintroductions, and the linkages
between these two pathways. Rectangles and hexagons represent outcomes
and processes, respectively. Blue text indicates components that are
included in the current study. We quantified frog survival and
population viability following reintroductions directly using field
studies and models, and evaluated ``disease-mediated selection''
indirectly using a comparison of frog exomes in naive and recovering
populations. The timeline shows the general sequence of the components,
with the dashed portion indicating a projection into the future.}
